---
title: "Project_1"
author: "Fabio_Savorgnan"
date: "1/30/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


__Library

```{r }
library(dplyr)
library(ggplot2)
library(tidyr)
#library(SDMTools)
library(readr)
library(digest)
library(ISLR)
library(car) 
library(leaps)
library( Matrix)
library(foreach)
library(glmnet)
library(VIM)
library(mice)
library(corrgram)
library(car)
library(tidyverse)
library(plyr)
#library(limma)
library(gridExtra)
```

## Load the data

```{r }
df<- read.csv(file.choose())
lifedata2014 = subset(df,Year==2014)
lifedata2014 = lifedata2014 %>% select(c(Country))
GDP = read.csv("C:/Users/BelajiAvvaru/Desktop/Docs/SMU/MSDS 6372/Project 1/GDP.csv", header=T, sep = ',')
str(GDP)
Population = read.csv("C:/Users/BelajiAvvaru/Desktop/Docs/SMU/MSDS 6372/Project 1/Population.csv", header=T, sep = ',')

lifedata2014GDP = merge(lifedata2014, GDP, by.x = "Country", by.y = "Country", all.x = TRUE)
summary(lifedata2014GDP)
head(df)
```





```{r }
#summary(df)
summary(df$Year)
summary(df$Life.expectancy)
summary(df$Adult.Mortality)
summary(df$infant.deaths)
summary(df$Alcohol)
summary(df$percentage.expenditure)
summary(df$Hepatitis.B)
summary(df$Measles)
summary(df$BMI)
summary(df$under.five.deaths)
summary(df$Polio)
summary(df$Total.expenditure)
summary(df$Diphtheria)
summary(df$HIV.AIDS)
summary(df$GDP)
summary(df$Population)
summary(df$Total.expenditure)
summary(df$thinness..1.19.years)
summary(df$thinness.5.9.years)
summary(df$Income.composition.of.resources)
summary(df$Schooling)
dim(df)
```

# Plot the data to perform DEA

```{r visual1, fig.height=2}
lifedata = subset(df,Year==2014)
summary(lifedata)
lifedatatest =lifedata %>% filter(lifedata$Measles > 1000)
lifedatatest1 =lifedata %>% filter(lifedata$under.five.deaths > 1000)
hist(lifedata$under.five.deaths)
 
```


# let's add another cateogorical varible for ANOVA analysis

```{r}

#Add new categorical variable "SchoolingLevel"
lifedata$SchoolingLevel<-cut(lifedata$Schooling, seq(4,24,5), right=FALSE, labels=c("1","2", "3", "4"))


```


# plot life expactancy itself

```{r}

ggplot(data=lifedata, aes(lifedata$Life.expectancy)) + 
  geom_histogram(binwidth = 1) + ggtitle("Histogram of Life Expectancy") + xlab("Life Expectancy")

```

# compare categorical variable vs life expectancy

```{r}
# histogram for life expectancy by status


mu <- ddply(lifedata, "Status", summarise, grp.mean=mean(lifedata$Life.expectancy))
#head(mu)




p <- ggplot(lifedata, aes(x=lifedata$Life.expectancy, fill=lifedata$Status, color=lifedata$Status)) +
  geom_histogram(position="identity")
# Use semi-transparent fill
p <- ggplot(lifedata, aes(x=lifedata$Life.expectancy, fill=lifedata$Status, color=lifedata$Status)) +
  geom_histogram(position="identity", alpha=0.5)

# Add mean lines
p <- p+geom_vline(data=mu, aes(xintercept=grp.mean, color=Status),
             linetype="dashed")

p

t(aggregate(lifedata$Life.expectancy~lifedata$Status,data=lifedata,summary))

```

#historgram vs school Level

```{r}
# histogram for life expectancy by status


mu <- ddply(lifedata, "SchoolingLevel", summarise, grp.mean=mean(lifedata$Life.expectancy))
#head(mu)




p <- ggplot(lifedata, aes(x=lifedata$Life.expectancy, fill=lifedata$SchoolingLevel, color=lifedata$SchoolingLevel)) +
  geom_histogram(position="identity")
# Use semi-transparent fill
p <- ggplot(lifedata, aes(x=lifedata$Life.expectancy, fill=lifedata$SchoolingLevel, color=lifedata$SchoolingLevel)) +
  geom_histogram(position="identity", alpha=0.5)

# Add mean lines
p <- p+geom_vline(data=mu, aes(xintercept=grp.mean, color=SchoolingLevel),
             linetype="dashed")

p

t(aggregate(lifedata$Life.expectancy~lifedata$SchoolingLevel,data=lifedata,summary))

```



# scatter plots of life expectancy vs. various data

```{r}

ggplot(lifedata, aes(x=lifedata$Adult.Mortality, y=lifedata$Life.expectancy)) + geom_point(data=lifedata) + geom_smooth(method=lm) +
  xlab("Adult Mortality") + ylab("Life Expectancy") + ggtitle("Adult Mortality vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$infant.deaths, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("Infact Deaths") + ylab("Life Expectancy") + ggtitle("Infant Deaths vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Alcohol, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("Alcohol") + ylab("Life Expectancy") + ggtitle("Alcohol vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Hepatitis.B, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("Hebapitis B") + ylab("Life Expectancy") + ggtitle("Hepatitis B  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Measles, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("Measles") + ylab("Life Expectancy") + ggtitle("Measles  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$BMI, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("BMI") + ylab("Life Expectancy") + ggtitle("BMI  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$BMI, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=loess) +
  xlab("BMI") + ylab("Life Expectancy") + ggtitle("BMI  vs Life Expectancy")


ggplot(lifedata, aes(x=lifedata$under.five.deaths, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("Under Five Deaths") + ylab("Life Expectancy") + ggtitle("Under Five Deaths  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Polio, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("Polio") + ylab("Life Expectancy") + ggtitle("Polio  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Total.expenditure, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("Total Expenditure") + ylab("Life Expectancy") + ggtitle("Total Expenditure  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Diphtheria, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("Diptheria") + ylab("Life Expectancy") + ggtitle("Diptheria  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$HIV.AIDS, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("HIV AIDs") + ylab("Life Expectancy") + ggtitle("HIV AIDs  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$GDP, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("GDP") + ylab("Life Expectancy") + ggtitle("GDP vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Population, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("Population") + ylab("Life Expectancy") + ggtitle("Population vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$thinness..1.19.years, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("Thinness 1 to 19 Years") + ylab("Life Expectancy") + ggtitle("Thinness 1 to 19 Years vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$thinness..1.19.years, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=loess) +
  xlab("Thinness 1 to 19 Years") + ylab("Life Expectancy") + ggtitle("Thinness 1 to 19 Years vs Life Expectancy")


ggplot(lifedata, aes(x=lifedata$thinness.5.9.years, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("Thinness 5 to 9 Years") + ylab("Life Expectancy") + ggtitle("Thinness 5 to 9 Years vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$thinness.5.9.years, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=loess) +
  xlab("Thinness 5 to 9 Years") + ylab("Life Expectancy") + ggtitle("Thinness 5 to 9 Years vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Income.composition.of.resources, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("Income Composition of Resources") + ylab("Life Expectancy") + ggtitle("Income Composition of Resources vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Schooling, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("Schooling") + ylab("Life Expectancy") + ggtitle("Schooling  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$percentage.expenditure, y=lifedata$Life.expectancy)) + geom_point() + geom_smooth(method=lm) +
  xlab("Percentage Expenditure") + ylab("Life Expectancy") + ggtitle("Percentage Expenditure  vs Life Expectancy")





```



# redo the plots, but this time include the SchoolingLevel as a grouping/color

```{r}
ggplot(lifedata, aes(x=lifedata$Adult.Mortality, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point(data=lifedata) + geom_smooth(method=lm) +
  xlab("Adult Mortality") + ylab("Life Expectancy") + ggtitle("Adult Mortality vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$infant.deaths, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("Infact Deaths") + ylab("Life Expectancy") + ggtitle("Infant Deaths vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Alcohol, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("Alcohol") + ylab("Life Expectancy") + ggtitle("Alcohol vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Hepatitis.B, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("Hebapitis B") + ylab("Life Expectancy") + ggtitle("Hepatitis B  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Measles, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("Measles") + ylab("Life Expectancy") + ggtitle("Measles  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$BMI, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("BMI") + ylab("Life Expectancy") + ggtitle("BMI  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$BMI, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=loess) +
  xlab("BMI") + ylab("Life Expectancy") + ggtitle("BMI  vs Life Expectancy")


ggplot(lifedata, aes(x=lifedata$under.five.deaths, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("Under Five Deaths") + ylab("Life Expectancy") + ggtitle("Under Five Deaths  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Polio, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("Polio") + ylab("Life Expectancy") + ggtitle("Polio  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Total.expenditure, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("Total Expenditure") + ylab("Life Expectancy") + ggtitle("Total Expenditure  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Diphtheria, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("Diptheria") + ylab("Life Expectancy") + ggtitle("Diptheria  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$HIV.AIDS, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("HIV AIDs") + ylab("Life Expectancy") + ggtitle("HIV AIDs  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$GDP, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("GDP") + ylab("Life Expectancy") + ggtitle("GDP vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Population, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("Population") + ylab("Life Expectancy") + ggtitle("Population vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$thinness..1.19.years, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("Thinness 1 to 19 Years") + ylab("Life Expectancy") + ggtitle("Thinness 1 to 19 Years vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$thinness..1.19.years, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=loess) +
  xlab("Thinness 1 to 19 Years") + ylab("Life Expectancy") + ggtitle("Thinness 1 to 19 Years vs Life Expectancy")


ggplot(lifedata, aes(x=lifedata$thinness.5.9.years, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("Thinness 5 to 9 Years") + ylab("Life Expectancy") + ggtitle("Thinness 5 to 9 Years vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$thinness.5.9.years, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=loess) +
  xlab("Thinness 5 to 9 Years") + ylab("Life Expectancy") + ggtitle("Thinness 5 to 9 Years vs Life Expectancy")


ggplot(lifedata, aes(x=lifedata$Income.composition.of.resources, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("Income Composition of Resources") + ylab("Life Expectancy") + ggtitle("Income Composition of Resources vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Schooling, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("Schooling") + ylab("Life Expectancy") + ggtitle("Schooling  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$percentage.expenditure, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point() + geom_smooth(method=lm) +
  xlab("Percentage Expenditure") + ylab("Life Expectancy") + ggtitle("Percentage Expenditure  vs Life Expectancy")




```



# what countries are on this separate curve on life expectancy - very low adult mortality, and low life expectancy

```{r}

p <-  ggplot(lifedata, aes(x=lifedata$Adult.Mortality, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point(data=lifedata) + geom_smooth(method=lm) +
  xlab("Adult Mortality") + ylab("Life Expectancy") + ggtitle("Adult Mortality vs Life Expectancy") +
geom_label(
    label=lifedata$Country,
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T
  )
p

ggsave("Mortality vs Life Expectancy.jpg", plot = p, width = 20, height = 20, units = "in")



p <- ggplot(lifedata, aes(x=lifedata$thinness..1.19.years, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=loess) +
  xlab("Thinness 1 to 19 Years") + ylab("Life Expectancy") + ggtitle("Thinness 1 to 19 Years vs Life Expectancy")+
geom_label(
    label=lifedata$Country,
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T
  )
p
ggsave("Thinness1.19 vs Life Expectancy.jpg", plot = p, width = 20, height = 20, units = "in")



```



# redo the plots, but this time include the Status (developed, developing) as a grouping/color

```{r}
ggplot(lifedata, aes(x=lifedata$Adult.Mortality, y=lifedata$Life.expectancy, color=lifedata$SchoolingLevel)) + geom_point(data=lifedata) + geom_smooth(method=lm) +
  xlab("Adult Mortality") + ylab("Life Expectancy") + ggtitle("Adult Mortality vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$infant.deaths, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("Infact Deaths") + ylab("Life Expectancy") + ggtitle("Infant Deaths vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Alcohol, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("Alcohol") + ylab("Life Expectancy") + ggtitle("Alcohol vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Hepatitis.B, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("Hebapitis B") + ylab("Life Expectancy") + ggtitle("Hepatitis B  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Measles, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("Measles") + ylab("Life Expectancy") + ggtitle("Measles  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$BMI, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("BMI") + ylab("Life Expectancy") + ggtitle("BMI  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$BMI, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=loess) +
  xlab("BMI") + ylab("Life Expectancy") + ggtitle("BMI  vs Life Expectancy")


ggplot(lifedata, aes(x=lifedata$under.five.deaths, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("Under Five Deaths") + ylab("Life Expectancy") + ggtitle("Under Five Deaths  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Polio, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("Polio") + ylab("Life Expectancy") + ggtitle("Polio  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Total.expenditure, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("Total Expenditure") + ylab("Life Expectancy") + ggtitle("Total Expenditure  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Diphtheria, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("Diptheria") + ylab("Life Expectancy") + ggtitle("Diptheria  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$HIV.AIDS, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("HIV AIDs") + ylab("Life Expectancy") + ggtitle("HIV AIDs  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$GDP, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("GDP") + ylab("Life Expectancy") + ggtitle("GDP vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Population, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("Population") + ylab("Life Expectancy") + ggtitle("Population vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$thinness..1.19.years, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("Thinness 1 to 19 Years") + ylab("Life Expectancy") + ggtitle("Thinness 1 to 19 Years vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$thinness..1.19.years, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=loess) +
  xlab("Thinness 1 to 19 Years") + ylab("Life Expectancy") + ggtitle("Thinness 1 to 19 Years vs Life Expectancy")


ggplot(lifedata, aes(x=lifedata$thinness.5.9.years, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("Thinness 5 to 9 Years") + ylab("Life Expectancy") + ggtitle("Thinness 5 to 9 Years vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$thinness.5.9.years, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=loess) +
  xlab("Thinness 5 to 9 Years") + ylab("Life Expectancy") + ggtitle("Thinness 5 to 9 Years vs Life Expectancy")


ggplot(lifedata, aes(x=lifedata$Income.composition.of.resources, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("Income Composition of Resources") + ylab("Life Expectancy") + ggtitle("Income Composition of Resources vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$Schooling, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("Schooling") + ylab("Life Expectancy") + ggtitle("Schooling  vs Life Expectancy")

ggplot(lifedata, aes(x=lifedata$percentage.expenditure, y=lifedata$Life.expectancy, color=lifedata$Status)) + geom_point() + geom_smooth(method=lm) +
  xlab("Percentage Expenditure") + ylab("Life Expectancy") + ggtitle("Percentage Expenditure  vs Life Expectancy")




```



```{r }

#df_filtered <- df %>% filter(Year == 2015 | Year == 2014)
#head(df_filtered)

```





```{r }
pairs(lifedata[,-c(1,2,3)])

lifedata

pairs(lifedata[,c(4,5,6,7,8,9,10)], col=lifedata$Status)
pairs(lifedata[,c(4,11,12,13,14,15,16,17)], col=lifedata$Status)
pairs(lifedata[,c(4,18,19,20,21,22)], col=lifedata$Status)


```



# Clean the data 

```{r }
lifedata = subset(df,Year==2014)
#lifedata <-df
# Verify any missing values in any variable
lifedata[!complete.cases(lifedata), ]

#Display missing-data patterns
md.pattern(lifedata, plot=TRUE, rotate.names = TRUE)


#Display missing-data in a bar-plot
mice_plot <- aggr(lifedata, col=c('navyblue','yellow'),
                  numbers=TRUE, sortVars=TRUE,
                  labels=names(lifedata), cex.axis=.7,
                  gap=3, ylab=c("Missing data","Pattern"))

#Data frame with all continuous variables 
#lifedata1 = lifedata %>% select(c(Country, Status, Life.expectancy, Adult.Mortality, Alcohol, Hepatitis.B, BMI, Polio, 
#                                        Total.expenditure, Diphtheria, GDP,Population, Income.composition.of.resources, 
#                                        thinness..1.19.years , thinness.5.9.years, Schooling, HIV.AIDS, under.five.deaths,
#                                        Measles, percentage.expenditure, infant.deaths)) 

lifedata1 = lifedata %>% select(-c(Year)) 

#correlation plot for all continuous variables
corrgram(lifedata1, order=TRUE,
         upper.panel=panel.cor, lower.panel=panel.pie, main="Life expectancy correlogram")

#Impute missing data with "CART"(Classification and Regression Trees) method
imputed_Data2 <- mice(lifedata1, m=5, maxit =5, method = 'cart', seed = 500)

#Data frame after imputation
completeData1 <- mice::complete(imputed_Data2,)
str(completeData1)

#we readd this variable since it's based on schooliing, not really it's own varible
completeData1$SchoolingLevel<-cut(completeData1$Schooling, seq(4,24,5), right=FALSE, labels=c("1","2", "3", "4"))

#Display missing-data in a bar-plot after imputation
mice_plot <- aggr(completeData1, col=c('navyblue','red'),
                  numbers=TRUE, sortVars=TRUE,
                  labels=names(completeData1), cex.axis=.7,
                  gap=3, ylab=c("Missing data","Pattern"))


```

#Create a regression model to check VIF
```{r }
#str(completeData2)
completeData2 = completeData1 %>% select(-c(Country, Status, SchoolingLevel)) 
life.model1<-lm(Life.expectancy~Adult.Mortality+infant.deaths+Alcohol+percentage.expenditure+Hepatitis.B+Measles+
            BMI+under.five.deaths+Polio+Total.expenditure+Diphtheria+HIV.AIDS+GDP+Population+
            thinness..1.19.years+thinness.5.9.years+Income.composition.of.resources+Schooling , data=completeData2)

summary(life.model1)
vif(life.model1)

```

## Below variables are statistically significant 
Adult.Mortality
infant.deaths
under.five.deaths
Total.expenditure
HIV.AIDS 
Income.composition.of.resources

## Below variables are colliniar variables
infant.deaths
under.five.deaths
thinness..1.19.years
thinness.5.9.years

Variables  "infant.deaths" and "under.five.deaths" are strongly correlated with each other with correlation of 1. variable "under.five.deaths has a correlation of -0.26 with Life.expectancy where as variable "infant.deaths" has a correlation of -0.24. We can keep "under.five.deaths" and remove "infant.deaths" from the model

Variables  "thinness..1.19.years" and "thinness.5.9.years" are strongly correlated with each other with correlation of 0.95. We can remove one of the variable from the model. Variable "thinness.5.9.years" has a correlation of -0.49 with Life.expectancy where as  "thinness..1.19.years" has a correlation of -0.46. We can keep thinness.5.9.years and remove thinness..1.19.years from the model

```{r }
#model without variables infant.deaths and thinness..1.19.years
life.model2<-lm(Life.expectancy~Adult.Mortality+Alcohol+percentage.expenditure+Hepatitis.B+Measles+
            BMI+under.five.deaths+Polio+Total.expenditure+Diphtheria+HIV.AIDS+GDP+Population+
            thinness.5.9.years+Income.composition.of.resources+Schooling , data=completeData2)

summary(life.model2)
vif(life.model2)

```

## Below variables are statistically significant 
Adult.Mortality
Total.expenditure
HIV.AIDS 
Income.composition.of.resources

## Multicolliniarity issue fixed after removing couple of variables from the model 

# OLS model with above signifcant variables 

```{r }
#model without variables infant.deaths and thinness..1.19.years
OLS.model<-lm(Life.expectancy~Adult.Mortality+Total.expenditure+HIV.AIDS+Income.composition.of.resources, data=completeData2)

summary(OLS.model)
plot(OLS.model)
confint(OLS.model)
vif(OLS.model)
```

# ASE with OLS

```{r }
OLStrain <- completeData2 [1:120, ]
OLStest <-completeData2 [121:183, ]
OLSModel1<-lm(Life.expectancy~Adult.Mortality+Total.expenditure+HIV.AIDS+Income.composition.of.resources, data=completeData2)
pred<-predict(OLSModel1, newdata=OLStest)
ASE<-mean((OLStest$Life.expectancy-pred)^2)
ASE

```

# Model selection "Forward"

```{r }
completeData2 = completeData1 %>% select(-c(Country, Status, SchoolingLevel)) 
reg.fwd=regsubsets(Life.expectancy~., data=completeData2, method="forward", nvmax=20)

reg.fwd
```



# Plot ASE

```{r }
set.seed(1234)
index<-sample(1:dim(completeData2)[1], 100, replace=F)
fwdtrain<-completeData2[index,]
fwdtest<-completeData2[-index,]
reg.fwd=regsubsets(Life.expectancy~., data=completeData2, method="forward", nvmax=15)

reg.fwd
predict.regsubsets =function (object , newdata ,id ,...){
  form=as.formula (object$call [[2]])
  mat=model.matrix(form ,newdata )
  coefi=coef(object ,id=id)
  xvars=names(coefi)
  mat[,xvars]%*%coefi
}

testASE<-c()
#note my index is to 20 since that what I set it in regsubsets
for (i in 1:15){
  predictions<-predict.regsubsets(object=reg.fwd,newdata=fwdtest,id=i) 
  testASE[i]<-mean((fwdtest$Life.expectancy-predictions)^2)
}

par(mfrow=c(1,1))
plot(1:15,testASE,type="l",xlab="# of predictors",ylab="test vs train ASE", ylim=c(5, 30))
index<-which(testASE==min(testASE))
points(index,testASE[index],col="red",pch=10)
rss<-summary(reg.fwd)$rss
lines(1:15,rss/100,lty=3,col="blue") 
```

### BIC, adj R-squared and RSS plots for forward selection

```{r }
par(mfrow=c(1,3))
bics<-summary(reg.fwd)$bic
plot(1:15,bics,type="l",ylab="BIC",xlab="# of predictors")
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)

adjr2<-summary(reg.fwd)$adjr2
plot(1:15,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)

rss<-summary(reg.fwd)$rss
plot(1:15,rss,type="l",ylab="train RSS",xlab="# of predictors")
index<-which(rss==min(rss))
points(index,rss[index],col="red",pch=10)
``` 


# Pick up the model
```{r }

m1<-coef(reg.fwd,4)
m1


```



# Model selection "stepwise"

```{r }
completeData2 = completeData1 %>% select(-c(Country, Status, SchoolingLevel)) 
reg.stp=regsubsets(Life.expectancy~., data=completeData2, method="seqrep", nvmax=20)

reg.stp
```



# Plot ASE

```{r }
set.seed(1234)
index<-sample(1:dim(completeData2)[1], 100, replace=F)
stptrain<-completeData2[index,]
stptest<-completeData2[-index,]
reg.stp=regsubsets(Life.expectancy~., data=completeData2, method="seqrep", nvmax=15)

reg.fwd
predict.regsubsets =function (object , newdata ,id ,...){
  form=as.formula (object$call [[2]])
  mat=model.matrix(form ,newdata )
  coefi=coef(object ,id=id)
  xvars=names(coefi)
  mat[,xvars]%*%coefi
}

stptestASE<-c()
#note my index is to 20 since that what I set it in regsubsets
for (i in 1:15){
  predictions<-predict.regsubsets(object=reg.stp,newdata=stptest,id=i) 
  stptestASE[i]<-mean((stptest$Life.expectancy-predictions)^2)
}

par(mfrow=c(1,1))
plot(1:15,stptestASE,type="l",xlab="iteration #",ylab="test vs train ASE", ylim=c(5, 50))
index<-which(stptestASE==min(stptestASE))
points(index,stptestASE[index],col="red",pch=10)
stprss<-summary(reg.stp)$rss
lines(1:15,stprss/100,lty=3,col="blue") 
```

### BIC, adj R-squared and RSS plots for stepwise selection

```{r }

par(mfrow=c(1,3))
bics<-summary(reg.stp)$bic
plot(1:15,bics,type="l",ylab="BIC",xlab="iteration #")
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)

adjr2<-summary(reg.stp)$adjr2
plot(1:15,adjr2,type="l",ylab="Adjusted R-squared",xlab="iteration #")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)

rss<-summary(reg.stp)$rss
plot(1:15,rss,type="l",ylab="train RSS",xlab="#iteration #")
index<-which(rss==min(rss))
points(index,rss[index],col="red",pch=10)
``` 


########################################################################################################################################################################################################################################################################



```{r }

ms<-coef(reg.stp,4)
ms

plot(reg.stp)

```

# LASSO



```{r,echo=T}
set.seed(1234)
index<-sample(1:dim(completeData2)[1],100,replace=F)
train<-completeData2[index,]
test<-completeData2[-index,]
#reg.fwd=regsubsets(log(AvgWinnings)~.,data=train,method="forward",nvmax=20)

```





```{r }
require("glmnet")
#Formatting data for GLM net
x=model.matrix(Life.expectancy~Adult.Mortality+Hepatitis.B+BMI+I(BMI^2)+I(BMI^3)+Polio+Total.expenditure+Diphtheria+GDP+Population
           +Income.composition.of.resources+thinness..1.19.years + I(thinness..1.19.years^2)+thinness.5.9.years+Schooling+HIV.AIDS
           +percentage.expenditure,train)[,-1]
y=train$Life.expectancy

xtest<-model.matrix(Life.expectancy~Adult.Mortality+Hepatitis.B+BMI+I(BMI^2)+I(BMI^3)+Polio+Total.expenditure+Diphtheria+GDP+Population
           +Income.composition.of.resources+thinness..1.19.years + I(thinness..1.19.years^2)+thinness.5.9.years+Schooling+HIV.AIDS
           +percentage.expenditure,test)[,-1]
ytest<-test$Life.expectancy



grid=10^seq(10,-2, length =100)
lasso.mod=glmnet(x,y,alpha=1, lambda =grid)

cv.out=cv.glmnet(x,y,alpha=1) #alpha=1 performs LASSO
plot(cv.out)
bestlambda<-cv.out$lambda.min  #Optimal penalty parameter.  You can make this call visually.
lasso.pred=predict (lasso.mod ,s=bestlambda ,newx=xtest)

testMSE_LASSO<-mean((ytest-lasso.pred)^2)
testMSE_LASSO



```



```{r}
coef(lasso.mod,s=bestlambda)
```



######################################################################################################################################################################## 2-Way ANOVA ###################################################################################


```{r }
#Run the lasso model
lasso.model<-lm(Life.expectancy~Adult.Mortality+Total.expenditure+HIV.AIDS+Income.composition.of.resources+ BMI + Diphtheria + Population +thinness..1.19.years + Hepatitis.B, data=completeData2)

summary(lasso.model)
plot(lasso.model)
confint(lasso.model)
vif(lasso.model)
```



# Test the lasso model
```{r }
lassotrain <- completeData2 [1:120, ]
lassotest <-completeData2 [121:183, ]

lasso.model<-lm(Life.expectancy~Adult.Mortality+Total.expenditure+HIV.AIDS+Income.composition.of.resources+ BMI + Diphtheria + Population +thinness..1.19.years + I(thinness..1.19.years^2) + Hepatitis.B, data=completeData2)

pred<-predict(lasso.model, newdata=lassotest)
ASE<-mean((lassotest$Life.expectancy-pred)^2)
ASE

```


# Coding the categorical variable for the 2 way anova

```{r visual1, fig.height=2}

lifedata = subset(df,Year==2014)
lifedata <-df
lifedata$SchoolingLevel<-cut(lifedata$Schooling, seq(5,30,5), right=FALSE, labels=c(1:6))
par(mfrow=c(1,3))

```




# Prepare data for 2-way ANOVA
```{r}
lifeanovadata = completeData1 %>% select(c(Status, SchoolingLevel, Life.expectancy)) 
str(lifeanovadata)
```

# compute Mean SD
```{r}

mysummary1<-function(x){
  result<-c(length(x),mean(x),sd(x))
  names(result)<-c("N","Mean","SD")
  return(result)
}
sumstats1<-aggregate(Life.expectancy~SchoolingLevel*Status,data=lifeanovadata,mysummary1)
sumstats1<-cbind(sumstats1[,1:2],sumstats1[,-(1:2)])
sumstats1
```

# Means plot using the standard deviations (SD) 

```{r}
ggplot(sumstats1,aes(x=SchoolingLevel,y=Mean,group=Status,colour=Status))+
  ylab("Life expectancy")+
  geom_line()+
  geom_point()+
  geom_errorbar(aes(ymin=Mean-SD,ymax=Mean+SD),width=.1)
``` 

# ANOVA model

```{r}
model.fit<-aov(Life.expectancy~Status+SchoolingLevel+Status:SchoolingLevel,data=lifeanovadata)
summary(model.fit)
par(mfrow=c(1,2))
plot(model.fit$fitted.values,model.fit$residuals,ylab="Resdiduals",xlab="Fitted")
qqnorm(model.fit$residuals)
```



# I different dysplay of the plot for the 2 way anova

```{r}
myfits<-data.frame(fitted.values=model.fit$fitted.values,residuals=model.fit$residuals)

#Residual vs Fitted
plot1<-ggplot(myfits,aes(x=fitted.values,y=residuals))+ylab("Residuals")+
  xlab("Predicted")+geom_point()

#QQ plot of residuals  #Note the diagonal abline is only good for qqplots of normal data.
plot2<-ggplot(myfits,aes(sample=residuals))+
  stat_qq()+geom_abline(intercept=mean(myfits$residuals), slope = sd(myfits$residuals))

#Histogram of residuals
plot3<-ggplot(myfits, aes(x=residuals)) + 
  geom_histogram(aes(y=..density..),binwidth=1,color="black", fill="gray")+
  geom_density(alpha=.1, fill="red")

grid.arrange(plot1, plot2,plot3, ncol=3)


```






```{r}

model.fit1<-aov(Life.expectancy~Status+SchoolingLevel,data=lifeanovadata)
Anova(model.fit1,type=3)
```

```{r}
library(car)
TukeyHSD(model.fit1,"Status",conf.level=.95)
TukeyHSD(model.fit1,"SchoolingLevel",conf.level=.95)
```